
# Chapter 5 - Dimensionality reduction techniques


* Part 1 - the "Human" data

We will start working by opening the data that has been wrangleg before. The dataset has 155 observations of 8 variables. It considers education, life expectation, labour force participation and so on in different countries in the world.


```{r, echo=FALSE}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", header = T, sep=",")

dim(human)
str(human)
```


```{r}
library(ggplot2)
library(GGally)

ggpairs(human)
summary(human)

```


Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them


Next we will perform a principal component analysis.

```{r}
# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

```

In the biplot nearly all of the observations are in the corner on the right. The plot is hard to interpret because the arrows and observations cannot be seen clearly. Better results will probably occur after standardising the data, so let's do the standardization next.

```{r}

human_std <- scale(human)

```



Next we will perform a principal component analysis with the standardized 'human' data.

```{r}
# perform principal component analysis (with the SVD method)
pca_human2 <- prcomp(human_std)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human2, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

```

This biplot is much more informatic, as the observations are clearer and the arrows now show the relationships between the variables in the dataset.

Interpret the results of both analysis (with and without standardizing). Are the results different? Why or why not? Include captions in you plots where you describe the results by using not just your variable names, but the actual phenomenons they relate to.



Give your personal interpretations of the first two principal component dimensions based on the biplot drawn after PCA on the standardized human data


-------

* Part 2 - the "tea" data from package FactoMineR


Let's first install the package and data in question.

```{r}
library(FactoMineR)

data("tea")

```


```{r}
dim(tea)
str(tea)
```

There are 300 observations of 36 variables in the data and it describes tea drinking.

visualize it.

We will now perform a Multiple Correspondence Analysis on the tea data (or to a certain columns of the data, it's up to you).

Interpret the results of the MCA

and draw at least the variable biplot of the analysis. You can also explore other plotting options for MCA.

Comment on the output of the plots










